\documentclass[10pt,twocolumn]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{times}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    citecolor=black
}

\setlength{\columnsep}{0.25in}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

\title{\textbf{Proof of Time: A Peer-to-Peer Electronic Cash System\\Based on Time}}
\author{
Alejandro Montana\\
\texttt{alejandromontana@tutamail.com}
}
\date{December 2025 \quad v2.0}

\begin{document}

\maketitle

\begin{abstract}
A purely peer-to-peer version of electronic cash where consensus is based on time, not capital. The main problem with existing systems is that influence is proportional to money: in Proof of Work --- to ASIC purchases, in Proof of Stake --- to token holdings. We propose a system where the only resource that matters is time. Time cannot be bought, accelerated, or transferred. Run a node for 180 days --- gain maximum influence. Your capital is irrelevant.
\end{abstract}

\section{Introduction}

Commerce on the Internet has evolved from trusted third parties to trustless consensus. Bitcoin solved the double-spending problem through Proof of Work. Ethereum added programmability through Proof of Stake. Both systems share a fundamental flaw: influence is bought with money.

In PoW, those who can afford more ASICs control the network. In PoS, those who hold more tokens control the network. The promise of decentralization collapses into plutocracy.

We need a system where:
\begin{itemize}[noitemsep]
    \item Influence cannot be purchased
    \item Entry barrier is zero
    \item Attack cost scales with time, not money
\end{itemize}

The solution is Proof of Time.

\section{Time as Consensus}

\subsection{The Problem with Capital-Based Consensus}

\begin{align*}
\text{PoW:} \quad & \text{Influence} = f(\text{Money} \rightarrow \text{Hardware} \rightarrow \text{Electricity}) \\
\text{PoS:} \quad & \text{Influence} = f(\text{Money} \rightarrow \text{Tokens}) \\
\text{PoT:} \quad & \text{Influence} = f(\text{Time})
\end{align*}

Time is the only resource distributed equally to all humans. One second for a billionaire equals one second for anyone else.

\subsection{Verifiable Delay Functions}

A Verifiable Delay Function (VDF) is a function that:
\begin{enumerate}[noitemsep]
    \item Requires sequential computation (cannot be parallelized)
    \item Produces a proof that can be quickly verified
    \item Has deterministic output
\end{enumerate}

We use Wesolowski VDF:
\begin{align}
y &= x^{2^T} \mod N \\
\pi &= x^{\lfloor 2^T / l \rfloor} \mod N
\end{align}

where $T$ = number of iterations (time parameter), $N$ = RSA modulus (2048-bit), $l$ = hash-derived prime.

Verification: $y = \pi^l \cdot x^r \mod N$ where $r = 2^T \mod l$

VDF guarantees that $T$ sequential squarings were performed. No amount of money can speed this up --- only time.

\section{Dual-Layer Architecture}

\subsection{Layer 1: Proof of History (PoH)}

Fast transaction layer. Sequential SHA-256 chain:
\begin{equation}
H(n) = \text{SHA256}(H(n-1) \| \text{data} \| \text{timestamp})
\end{equation}

\begin{itemize}[noitemsep]
    \item 1 hash per slot
    \item Transactions embedded in hash chain
    \item Provides ordering without consensus
\end{itemize}

\subsection{Layer 2: Proof of Time (PoT)}

Finality layer. VDF checkpoints every 600 seconds:
\begin{equation}
\text{Checkpoint}(n) = \text{VDF}(H(\text{last\_slot}), T=10^6)
\end{equation}

\begin{itemize}[noitemsep]
    \item Cannot be reverted after VDF completion
    \item 10-minute finality (like Bitcoin, but deterministic)
    \item Checkpoint = irreversible anchor
\end{itemize}

\section{Leader Selection}

\subsection{ECVRF}

Leader is selected via Elliptic Curve Verifiable Random Function:
\begin{align}
(\beta, \pi) &= \text{VRF\_prove}(sk, \text{seed}) \\
\text{VRF\_verify}(pk, \text{seed}, \beta, \pi) &\rightarrow \{0, 1\}
\end{align}

Where: $\text{seed}$ = previous checkpoint hash, $\beta$ = pseudo-random output, $\pi$ = proof of correct computation.

\subsection{Selection Probability}

\begin{equation}
P(\text{node}_i) = \frac{\text{Adonis}(i)}{\sum_j \text{Adonis}(j)}
\end{equation}

Selected if: $\beta < P(\text{node}_i) \times 2^{256}$

The Adonis score determines selection probability.

\section{The Six Dimensions of Adonis}

Node weight is computed from six dimensions, each measuring a different aspect of node contribution.

\subsection{Formula}

\begin{equation}
\text{Adonis}(i) = \sum_{d \in D} w_d \times \text{score}_d(i)
\end{equation}

\subsection{Dimensions}

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}llrl@{}}
\toprule
\textbf{Dimension} & \textbf{Domain} & \textbf{Weight} & \textbf{Saturation} \\
\midrule
TIME & Uptime & 35\% & 180 days \\
INTEGRITY & Behavior & 22\% & No violations \\
STORAGE & Data & 15\% & 100\% history \\
GEOGRAPHY & Location & 12\% & Diversity \\
RELIABILITY & Performance & 8\% & 99.9\% uptime \\
STAKE & Collateral & 8\% & Optional \\
\bottomrule
\end{tabular}
\end{table}

\subsection{TIME --- 35\%}

\begin{equation}
\text{score}_{\text{time}} = \min\left(\frac{\text{uptime\_seconds}}{15,552,000}, 1.0\right)
\end{equation}

15,552,000 seconds = 180 days. After 180 days, newcomer equals veteran.

\subsection{INTEGRITY --- 22\%}

Behavioral score. Positive actions increase, violations decrease:

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Event} & \textbf{Change} \\
\midrule
BLOCK\_PRODUCED & +0.05 \\
BLOCK\_VALIDATED & +0.02 \\
BLOCK\_INVALID & --0.15 \\
EQUIVOCATION & --1.0 + 180d quarantine \\
\bottomrule
\end{tabular}
\end{table}

\subsection{STORAGE --- 15\%}

\begin{equation}
\text{score}_{\text{storage}} = \min\left(\frac{\text{stored\_blocks}}{\text{total\_blocks}}, 1.0\right)
\end{equation}

Full nodes store complete history. Light nodes get proportional score.

\subsection{GEOGRAPHY --- 12\%}

Incentivizes global distribution:
\begin{equation}
\text{geo} = 0.6 \times \text{country\_rarity} + 0.4 \times \text{city\_rarity}
\end{equation}

First node from new country: +0.25 bonus. First node from new city: +0.15 bonus.

\subsection{RELIABILITY --- 8\%}

\begin{equation}
\text{score}_{\text{rel}} = \frac{\text{successful\_responses}}{\text{total\_requests}}
\end{equation}

Measures node availability and response quality.

\subsection{STAKE --- 8\%}

Optional collateral for enhanced trust. Not required for participation.

\section{DAG Structure}

\subsection{Block References}

Each block references 1--8 parent blocks:

\begin{verbatim}
Block {
  parents: [hash_1, ..., hash_k]
  transactions: [...]
  vrf_proof: pi
  timestamp: t
}
\end{verbatim}

\subsection{PHANTOM-PoT Ordering}

Blocks are ordered by:
\begin{enumerate}[noitemsep]
    \item VDF checkpoint anchors
    \item Topological sort within checkpoint window
    \item Tie-breaking via block hash
\end{enumerate}

Horizontal scaling: more parents = higher throughput.

\section{Economics}

\subsection{Unit}

\begin{equation}
1 \text{ Jot (\textit{\DJ})} = 1 \text{ second of time}
\end{equation}

\subsection{Emission}

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}ll@{}}
\toprule
Total supply & 1,260,000,000 \DJ \\
Block time & 10 minutes \\
Initial reward & 50 \DJ \\
Halving & Every 210,000 blocks \\
Full emission & $\sim$132 years \\
\bottomrule
\end{tabular}
\end{table}

Same curve as Bitcoin. Predictable, deflationary.

\section{Privacy Tiers}

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}lllr@{}}
\toprule
\textbf{Tier} & \textbf{Hidden} & \textbf{Size} & \textbf{Fee} \\
\midrule
T0 & Nothing & 250 B & 1$\times$ \\
T1 & Receiver & 400 B & 2$\times$ \\
T2 & + Amount & 1.2 KB & 5$\times$ \\
T3 & + Sender & 2.5 KB & 10$\times$ \\
\bottomrule
\end{tabular}
\end{table}

Privacy is optional. User chooses transparency vs. anonymity.

\section{Attack Analysis}

\subsection{Sybil Attack}

Creating $N$ fake nodes: each needs 180 days to reach TIME saturation. No shortcut.
\begin{equation}
\text{Attack cost} = N \times 180 \text{ days}
\end{equation}

\subsection{51\% Attack}

To control 51\% of Adonis weight with 1000 existing nodes:
\begin{equation}
\text{Cost} = 1020 \times 180 = 183,600 \text{ node-days}
\end{equation}

Compare:
\begin{itemize}[noitemsep]
    \item Bitcoin 51\%: $\sim$\$20B in hardware
    \item Ethereum 51\%: $\sim$\$10B in stake
    \item Proof of Time 51\%: $N \times 180$ days (cannot be bought)
\end{itemize}

\subsection{Long-Range Attack}

VDF checkpoints are irreversible. Rewriting history requires recomputing all VDFs from fork point. Each VDF takes real time. Honest chain always ahead. Not feasible.

\section{Comparison}

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}lccc@{}}
\toprule
& \textbf{Bitcoin} & \textbf{Ethereum} & \textbf{PoT} \\
\midrule
Consensus & PoW & PoS & VDF \\
Influence & Money & Money & Time \\
Entry cost & High & Medium & Zero \\
Energy & Massive & Low & Minimal \\
51\% cost & \$20B & \$10B & Time \\
Finality & Prob. & 15 min & 10 min \\
\bottomrule
\end{tabular}
\end{table}

\section{Conclusion}

We have proposed a system for electronic transactions that does not rely on capital for consensus. Time is the only resource that cannot be bought, accelerated, or transferred.

The network self-organizes through the Six Dimensions of Adonis:
\begin{itemize}[noitemsep]
    \item \textbf{TIME} ensures long-term commitment
    \item \textbf{INTEGRITY} removes bad actors
    \item \textbf{STORAGE} maintains data availability
    \item \textbf{GEOGRAPHY} enforces global distribution
    \item \textbf{RELIABILITY} rewards performance
    \item \textbf{STAKE} provides optional collateral
\end{itemize}

The result is a system where everyone starts equal, influence is earned not bought, attacks require time not money, and decentralization is incentivized not just promised.

\begin{center}
\textit{In time, we are all equal.}
\end{center}

\section*{References}

\begin{enumerate}[noitemsep]
    \item Nakamoto, S. (2008). Bitcoin: A Peer-to-Peer Electronic Cash System.
    \item Wesolowski, B. (2019). Efficient Verifiable Delay Functions.
    \item Boneh, D., et al. (2018). Verifiable Delay Functions.
    \item Sompolinsky, Y., Zohar, A. (2015). PHANTOM: A Scalable BlockDAG Protocol.
    \item Micali, S., Rabin, M., Vadhan, S. (1999). Verifiable Random Functions.
\end{enumerate}

\end{document}
